Cascades, Product Conversion, Whatever

### 2012-12-12 paper.tex
Finished future work, and corrected some stuff. Passed my initial test

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}

@@

We propose a cascade spreading model that borrows heavily from the
scheduler mediated cascade process introduced in
\cite{Chierichetti12}. In our model every node in an undirected graph
is a risk neutral rational agent. Each node is also assigned a binary
type which affects their utility, pulled from a binomial
distribution. In this paper we call the types Yes ($Y$) and No
($N$). There is an additional scheduler agent, which, at each round of
the game picks one of the nodes in the graph. At that moment the
chosen node chooses to commit to a choice (also Yes or No) and the
agent is committed to that choice for the rest of the game. The game
ends when all nodes have committed to a choice.

@@

The full game can be described by the following algorithm. The
scheduler picks an optimal first node in the graph. That node chooses
either Yes or No based on all of the parameters and its
knowledge. Based on this result the scheduler picks a new node, and so
on until every node is chosen and the game finishes.

Given this game, we are interested in quantifying
\begin{equation*}
  \frac{\mathbb E[\#Y|strategic]}{\mathbb E[\#Y|myopic]}.
\end{equation*}

@@

A graph is a $k$-block model is the vertices can be divided into $k$
disjoint subsets such that every node in the same subset has
incoming and outgoing edges to the same set of nodes. Another way to
view this property is by looking at the adjacency matrix of the
graph. A block model graph can be written as a symmetric block matrix
with $k^2$ blocks. For the sake of the algorithm, we will define
\begin{equation*}
  \{s_i\}_{i=1..k}
\end{equation*}
as the number of nodes in each block.

@@

The size of the node table is trivial to upper bound. There are $2k+2$
dimensions: the node's type (2 values), the node's block ($k$ values),
$k$ dimensions for the number of nodes in each block that have gone
before ($\le s_i + 1 | i=1..k$), and $k$ dimensions for the number of
nodes in each block that have gone before and selected yes ($\le s_i +
1 | i=1..k$). This leads to the following statement,
\begin{align*}
  NodeTableSize &= o\left( 2k \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= o\left( k \prod_{i=1}^k s_i^2 \right)
  \intertext{If we further assume that $\ell \le k$ of
    the blocks are size order $n$ and the rest are constant size, then the
    size of the table reduces to}
  NodeTableSize &= o\left( kn^{2\ell} \right)
  \intertext{Each node table entry
    computation takes 2 calls to the sched function, and then does
    computation of the expected number of Yeses which is order $k$,
    making the total complexity for the node table given the sched
    table}
  NodeTableComplexity &= o\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  NodeTableComplexity &= o\left( k^2 n^{2\ell} \right)
  \intertext{The size of the scheduler table is also trivial to upper
    bound. There are $2k$ dimensions similar to the node table. By
    similar reasoning}
  SchedTableSize &= o\left( \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= o\left( \prod_{i=1}^k s_i^2 \right)
  \intertext{If we make the same assumption about block sizes this
    reduces to}
  SchedTableSize &= o \left( n^{2\ell} \right)
  \intertext{The complexity for each entry in the scheduler table is
    linear in $k$, because it makes $2k$ lookups to the node
    table. Therefore the total complexity for constructing the
    scheduler table is bounded by}
  SchedTableComplexity &= o\left( k \prod_{i=1}^k s_i^2 \right) \\
  SchedTableComplexity &= o \left( k n^{2\ell} \right)
  \intertext{This leaves us with final time complexity of}
  RunningTime &= o\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  &= o\left( k^2 n^{2\ell} \right)
\end{align*}

@@

We test a star graph, a clique, and a special graph we call a cloud graph. A cloud graph consists of two singular vertices of degree one, one singular vertex of degree two, and two clouds of many vertices, each of degree two. Each of the singular degree one vertices is connected to every node in a distinct cloud. The singular degree two vertex is connected to every node in both clouds. This graph, under certain parameters, provably obtains more $Y$ adoptions for strategic users.

@@

The ratio is increasing in $p$ for the star graph, decreasing in $p$ for the clique, and both (depending on $\pi$) for the cloud graph. Even more strangely, the ratio is (mostly) increasing in $\pi$ for the star graph but depends on the $p$ value for the clique and cloud graphs. We see markedly different behavior when $\pi = 1$. This is the point at which agents are indifferent between choosing their own type and going with the expected majority. This indifference is unique to $\pi=1$, as our tie breaking methods lead to significant differences in outcome.

We learn several things regarding the question of whether strategic agents can ever give more $Y$ adoptions than myopic agents. Data for the star graph supports our theoretical result that myopic is always better on the star graph (and thus the ratio is bounded by 1). The clique plot suggests an interesting result. For small $p, \pi$, strategic is better than myopic. This is an unexpected, but explainable result. For $\pi=1+\epsilon$, only 1 $Y$ decision is required to start a $ƒ$ cascade with strategic users, but 2 $Y$ decisions are required with myopic users. So the ratio can be as large as $\frac{1}{p}$. We also confirm some cascade proofs we've had for the clique and cloud graphs, namely that we can achieve either a $Y$ or an $N$ cascade on the cloud graph, depending on the parameters, and that for many parameters the clique achieves an $N$ cascade.

@@

Talk about the significance, patterns, stuff like that that? Talk
about future work in quantifying for general graphs, although it may
not be possible?? At least in some nice form.

Patterns may not exist, but there may still be order level
relationships / bounds that are possible.

See if we can prove the problem is PPAD or NP-complete.

Offline scheduling

There are a number of possible extensions for this algorithm that
should also broaden the class of graphs that our results could apply
to. These extensions include scaling to $k$-block graphs, directed
graphs, and stochastic block graphs. Extending the algorithm to
$k$-block graphs shouldn’t conceptually change the algorithm, but we
believe that the original algorithm may scale exponentially in $k$ for
a $k$-block graph. However, for bounded $k$, the complexity is still
be moderately efficient. Additionally, scaling the algorithm up to
$k$-blocks may expose room for improvement in the algorithm and
provide further insight into quantifying the difference between myopic
and strategic agents in this setting. Modifying the algorithm to
account for directed graphs as long as they still fit into a
blockmodel is also likely trivial and would further increase the
number of graphs that we can analyze theoretically although the
applicability of directed graphs is not immediately apparent. Another
potentially easy extension of this would be to look at stochastic
blockmodels\cite{Wang87, Snijders97}. It seems reasonable that the
algorithm for static block graph models might extend in expectation to
unrealized (and hence offline scheduled) stochastic blockmodels. This
extension could likely never extend to realized graphs or to online
schedules, but stochastic blockmodels are much closer to graphs found
experimentally than $k$-block graphs.

>>>>

\usepackage{natbib}
\usepackage{array}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}

@@

We propose a cascade spreading model that borrows heavily from the
scheduler mediated cascade process introduced in
\cite{Chierichetti12}. In our model every node in an undirected graph
is a risk neutral rational agent. Each node is also assigned a binary
type which affects their utility, pulled from a binomial
distribution. In this paper we call the types Yes ($Y$) and No
($N$). There is an additional scheduler agent, which, at each round of
the game picks one of the nodes in the graph. The chosen node picks a
``choice'' (also Yes or No) and the agent is committed to that
``choice'' for the rest of the game. The game ends when all nodes have
picked a ``choice.''

@@

The full game can be described by the following algorithm. The
scheduler picks an optimal first node in the graph. That node chooses
either Yes or No based on all of the parameters and its
knowledge. Based on this result the scheduler picks a new node, and so
on until every node has been picked and the game finishes.

Given this game, we are interested in quantifying
\begin{equation*}
  \frac{\mathbb E[\#Y|strategic]}{\mathbb E[\#Y|myopic]}
\end{equation*}
for a given family of graphs with different game parameters.

@@

A graph is a $k$-block model if the vertices can be divided into $k$
disjoint subsets such that every node in the same subset has
incoming and outgoing edges to the same set of nodes. Another way to
view this property is by looking at the adjacency matrix of the
graph. A block model graph can be written as a symmetric block matrix
with $k^2$ blocks. For the sake of the algorithm, we will define
\begin{equation*}
  \{s_i\}_{i=1..k}
\end{equation*}
as the number of nodes in each block.

An example of a 2-block model would be the star graph, which can be
expressed in adjacency matrix form as
\begin{equation*}
  \begin{array}{r}
    1 \{ \\
    n-1 \{
  \end{array}
  \left[ \begin{array}{cc}
      0 & 1 \\
      1 & 0
      \end{array} \right]
\end{equation*}
and has $\mathbf s = \{1, n-1\}$.

@@

The size of the node table is trivial to upper bound. There are $2k+2$
dimensions: the node's type (2 values), the node's block ($k$ values),
$k$ dimensions for the number of nodes in each block that have gone
before ($\le s_i + 1 | i=1..k$), and $k$ dimensions for the number of
nodes in each block that have gone before and selected yes ($\le s_i +
1 | i=1..k$). This leads to the following statement,
\begin{align*}
  NodeTableSize &= o\left( 2k \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= o\left( k \prod_{i=1}^k s_i^2 \right)
  \intertext{If we further assume that $\ell \le k$ of
    the blocks are size order $n$ and the rest are constant size, then the
    size of the table reduces to}
  NodeTableSize &= o\left( kn^{2\ell} \right)
  \intertext{Each node table entry
    computation takes 2 calls to the sched function, and then does
    computation of the expected number of Yeses which is order $k$,
    making the total complexity for the node table given the sched
    table}
  NodeTableComplexity &= o\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  &= o\left( k^2 n^{2\ell} \right)
  \intertext{The size of the scheduler table is also trivial to upper
    bound. There are $2k$ dimensions similar to the node table. By
    similar reasoning}
  SchedTableSize &= o\left( \prod_{i=1}^k s_i+1 \prod_{i=1}^k s_i+1
  \right) \\
  &= o\left( \prod_{i=1}^k s_i^2 \right)
  \intertext{If we make the same assumption about block sizes this
    reduces to}
  SchedTableSize &= o \left( n^{2\ell} \right)
  \intertext{The complexity for each entry in the scheduler table is
    linear in $k$, because it makes $2k$ lookups to the node
    table. Therefore the total complexity for constructing the
    scheduler table is bounded by}
  SchedTableComplexity &= o\left( k \prod_{i=1}^k s_i^2 \right) \\
  &= o \left( k n^{2\ell} \right)
  \intertext{This leaves us with final time complexity of}
  RunningTime &= o\left( k^2 \prod_{i=1}^k s_i^2 \right) \\
  &= o\left( k^2 n^{2\ell} \right)
\end{align*}

@@

We test a star graph, a clique, and a special graph we call a cloud
graph. A cloud graph consists of two singular ``outer'' vertices of
degree $a$ and $b$ respectively, one singular ``inner'' vertex of
degree $a+b$, and two clouds of size $a$ and $b$ respectively, with
each vertex of degree two. Each of the ``outer'' vertices is connected
to every node in a distinct cloud. The singular ``inner'' vertex is
connected to every node in both clouds. This graph, under certain
parameters, provably obtains more $Y$ adoptions for strategic users.

@@

The ratio is increasing in $p$ for the star graph, decreasing in $p$
for the clique, and both (depending on $\pi$) for the cloud
graph. Even more strangely, the ratio is (mostly) increasing in $\pi$
for the star graph but depends on the $p$ value for the clique and
cloud graphs. We see markedly different behavior when $\pi = 1$. This
is the point at which agents are indifferent between choosing their
own type and going with the expected majority. This indifference is
unique to $\pi=1$, as our tie breaking methods lead to significant
differences in outcome (nodes pick No in the event of a tie).

We learn several things regarding the question of whether strategic
agents can ever give more $Y$ adoptions than myopic agents. Data for
the star graph supports our theoretical result that myopic is always
better on the star graph (and thus the ratio is bounded by 1). The
clique plot suggests an interesting result. For small $p, \pi$,
strategic is better than myopic. This is an unexpected, but
explainable result. For $\pi=1+\epsilon$, only 1 $Y$ decision is
required to start a $Y$ cascade with strategic users, but 2 $Y$ decisions
are required with myopic users. So the ratio can be as large as
$\frac{1}{p}$. We also confirm some cascade proofs we've had for the
clique and cloud graphs, namely that we can achieve either a $Y$ or an
$N$ cascade on the cloud graph, depending on the parameters, and that
for many parameters the clique achieves an $N$ cascade.

@@

Our results are interesting and certainly show that the difference
between optimal play with myopic and strategic users is a complex
space even very simple block graphs. However, there are still a number
of interesting unexplored questions that we have not looked at in the
scope of the problem.

Despite the fact that obvious patterns don't exist in the performance
ratio, it is still possible that the performance ratio can be bounded
in certain circumstance beyond the star graph. Proving bounds on the
performance ratio could help characterize the discontinuous regions in
the plots where myopic superiority switches in favor of strategic
agents.

We have only shown a polynomial time algorithm for graphs where the
number of blocks is constant with respect to $N$. A further and
important step would be to come up with an algorithm for general
graphs, or to reduce the problem to another complexity class like PPAD
or NP.

The algorithm as it is presented only works for online
scheduling. This is a necessary condition to allow the dynamic
programming to work, however ongoing research may suggest that for
offline scheduling, myopic agents will always permit a higher number
of Yes nodes. Looking at performance bounds under offline scheduling
could pose an viable avenue to study performance difference where
online schedules became very complicated. It is not apparent how to
compute an optimal offline schedule efficiently for a block graph or
otherwise.
